### 2025-02-21 products/manuscript/supplement/Supplementary-Material.qmd

---
title: "Supplement to Yufei Wu's MADA Project"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Overview

You can find all the supplemetary materials (figures and tables) in the supplement.

# Additional results

@tbl-sum_prev_wage shows a summary table for prevailing wage.

```{r}
#| label: tbl-sum_prev_wage
#| tbl-cap: "Summary Table for Prevailing Wage."
#| echo: FALSE
table1 = readRDS(here("results","tables","sum_prev_wage.rds"))
knitr::kable(table1)
```

@fig-bottom_wage_by_soc_title shows a bar plot of bottom 10 job titles by lowest median wage. @fig-bottom_wage_by_employer shows a bar plot of bottom 10 employers by lowest median wage. @fig-bottom_wage_by_state shows a bar plot of bottom 10 states by lowest median wage.

```{r}
#| label: fig-bottom_wage_by_soc_title
#| fig-cap: "Bottom 10 Job Titles by Lowest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","bottom_wage_by_soc_title.png"))
```

```{r}
#| label: fig-bottom_wage_by_employer
#| fig-cap: "Bottom 10 Employers by Lowest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","bottom_wage_by_employer.png"))
```

```{r}
#| label: fig-bottom_wage_by_state
#| fig-cap: "Bottom 10 States by Lowest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","bottom_wage_by_state.png"))
```

{{< pagebreak >}}

# Discussion

{{< pagebreak >}}