### 2025-03-14 products/manuscript/Manuscript.qmd
part 3 commit 1

---
title: "Analysis of H-1B Visa Workers Using H-1B LCA Disclosure Data (2020–2024)"
author: "Yufei Wu"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

* Yufei Wu$^{1}$ (ORCID: 0009-0003-5051-0623)

**Author affiliations**

1. Department of Chemistry, University of Georgia, Athens, GA, USA.

$\land$ Corresponding author: yufeiwu@uga.edu

{{< pagebreak >}}

# Summary/Abstract

{{< pagebreak >}}

# Introduction

## General Background Information

H-1B is a nonimmigrant visa category that allows employers to hire foreign workers in specialty occupations, like methematics, engineering, technology, and medical sciences [@H1BVisaProgram2016]. It requires the employee to have a bachelor's degree or equivalent in the specific specialty. Typically, the maximun duration of an H-1B visa is six years.It is the most common work visa in the US. There are 65,000 available H-1B visas each year, with 20,000 additional visas for candidates with a master's or doctorate degree from a U.S. institution. If there are more than 65,000 applications, USCIS will run a lottery to decide who can file an H-1B petition. As USCIS is receiving more and more H-1B registration these years, it's harder for a foreign worker to get an H-1B visa. The Labor Condition Application (LCA) is an application filed by employers to apply for work authorization on behalf of employees as a prerequisite for H-1B [@LaborConditionApplication2025]. The LCA contains essential details such as job title, wage, and location about the proposed H-1B employment. Therefore, analyzing the LCA data can provide some ideas on the H-1B application status.

## Description of data and data source

The data is H1B LCA Disclosure Data (2020-2024) from [Kaggle](https://www.kaggle.com/datasets/zongaobian/h1b-lca-disclosure-data-2020-2024/data). It contains LCA disclosure datasets from U.S. Department of Labor, covering the period from 2020 to 2024, and includes information such as case status, job title, Standard Occupational Classification (SOC) title, location, and wages.

## Questions/Hypotheses to be addressed

This analysis will primarily focus on certified cases. The key research questions to be examined include:

(1)Distribution Analysis:

What is the distribution of certified cases across various features such as SOC titles, locations, wages, and employer names?

(2)Variable Relationships:

How do wages vary across different states, SOC titles, and employers?

(3)Employment Trends and Remote Work Patterns:

How have employment patterns evolved over time?

What trends can be observed in remote work prevalence before and after the pandemic?

(4)Predictive Modeling:

Can a predictive model be developed to estimate wages based on job-related and employer-specific features?

Which factors contribute most significantly to wage determination in certified cases?

{{< pagebreak >}}

# Methods

## Data aquisition and cleaning

The dataset used in this analysis is the H1B LCA Disclosure Data (2020-2024) from [Kaggle](https://www.kaggle.com/datasets/zongaobian/h1b-lca-disclosure-data-2020-2024/data). It comprises labor condition application (LCA) disclosure records from the U.S. Department of Labor, covering the period from 2020 to 2024. The dataset includes key information such as case status, job title, Standard Occupational Classification (SOC) title, employer details, location, and wages.

To ensure data quality, the dataset was filtered to include only certified cases and those specifically related to H-1B visa applications.

Additionally, the dataset was refined to include only the following essential variables:

RECEIVED_DATE (date the application was received); SOC_TITLE (occupational title classified under the SOC system); EMPLOYER_NAME (name of the employer); EMPLOYER_STATE (employer’s location); WORKSITE_STATE (location of the worksite); WAGE_RATE_OF_PAY_FROM (wage offered to the worker); WAGE_UNIT_OF_PAY (unit of pay values); PREVAILING_WAGE (prevailing wage for the job); PW_UNIT_OF_PAY (unit of pay values)

Observations with missing values in the selected variables were removed. All character columns except state names were standardized to lowercase.

## Statistical analysis
Various plots were created to visualize the data. Classification models were used to analyze the categorical outcome.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

### Univariate Analysis: Distribution Analysis

The @fig-top10_soc_title displays the 10 job titles with the highest number of applications filed between 2020 and 2024, with "Software Developers, Applications" leading at over 500,000 cases. The @fig-top10_emp_name highlights the top 10 employers by application count, with "COGNIZANT TECHNOLOGY SOLUTIONS US CORP" topping the list at approximately 93,000 cases. The @fig-top10_emp_state presents the 10 states with the most applications, with California recording over 600,000 cases. Meanwhile, the @fig-distribution_PREV_WAGE illustrates the distribution of prevailing wages, revealing that most cases offer wages near $100,000 per year, while fewer cases fall at the lower end (around $15,000) or approach higher values near $800,000 (see Supplement Table 1).
```{r}
#| label: fig-top10_soc_title
#| fig-cap: "Top 10 Job Titles."
#| echo: FALSE
knitr::include_graphics(here("results","figures","top10_soc_title.png"))
```

```{r}
#| label: fig-top10_emp_name
#| fig-cap: "Top 10 Employers."
#| echo: FALSE
knitr::include_graphics(here("results","figures","top10_emp_name.png"))
```

```{r}
#| label: fig-top10_emp_state
#| fig-cap: "Top 10 Employer Locations."
#| echo: FALSE
knitr::include_graphics(here("results","figures","top10_emp_state.png"))
```

```{r}
#| label: fig-distribution_PREV_WAGE
#| fig-cap: "Distribution of Prevailing Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","distribution_PREV_WAGE.png"))
```

### Bivariate Analysis

The @fig-wage_by_soc_title shows the 10 job titles with the highest median wage, and they are all medical practitioners. The @fig-wage_by_employer highlights the 10 employers with the highest median wage, spanning healthcare, financial services, life sciences, technology, and manufacturing. The @fig-wage_by_state indicates that Washington has the highest median wage by state. Plots for the 10 lowest median wages are provided in Supplement Figures 1–3. The @fig-remote_trend_absolute illustrates the trend of remote work over the years. The number of remote work cases surged from 2019 to 2020 due to the pandemic and remained relatively stable from 2020 to 2023. Post-pandemic, there was a slight decline from 2023 to 2024. However, as remote work patterns became more established, the decrease was not significant. This analysis uses absolute values instead of proportions because remote work is identified based on discrepancies between employer and worksite states. Some remote work cases may not be captured under this definition, potentially underestimating the actual proportion. Therefore, absolute values provide a more accurate reflection of the remote work trend.

```{r}
#| label: fig-wage_by_soc_title
#| fig-cap: "Top 10 Job Titles by Highest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","wage_by_soc_title.png"))
```

```{r}
#| label: fig-wage_by_employer
#| fig-cap: "Top 10 Employers by Highest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","wage_by_employer.png"))
```

```{r}
#| label: fig-wage_by_state
#| fig-cap: "Top 10 States by Highest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","wage_by_state.png"))
```

```{r}
#| label: fig-remote_trend_absolute
#| fig-cap: "Remote Work Trends Over Years."
#| echo: FALSE
knitr::include_graphics(here("results","figures","remote_trend_absolute.png"))
```

## Basic statistical analysis
### Bivariate Analysis
The variable WAGE_RATE_OF_PAY_FROM was selected as the outcome of interest to study the wages of H-1B workers. I first conducted a bivariate analysis to identify potential predictor variables. To perform the analysis, I selected a half-month subset of the data.

The @fig-distribution_WAGE_Jan2024 illustrates a histogram of wage distribution during January 1–15, 2024. The continuous wage variable was categorized into four ranges: "(0, 75,000)", "(75,000, 150,000)", "(150,000, 225,000)", and "Above 225,000", converting it into a categorical variable.

A random forest model was applied for classification using a single predictor, with 1,380 observations in the training set and 346 observations in the testing set. @tbl-combined_accuracy1 to @tbl-combined_accuracy4 summarize the model fit for the predictors: job title (SOC_TITLE), employer location (EMPLOYER_STATE), worksite location (WORKSITE_STATE), and employer (EMPLOYER_NAME). The results indicate that these variables exhibit moderate predictive performance.

Next, a multivariate random forest model was developed. @tbl-combined_accuracy_all shows that the multivariate model outperformed the bivariate models.

In the next step, different classification algorithms will be explored to further enhance model performance. Additionally, cross-validation will be incorporated to improve model reliability.
```{r}
#| label: fig-distribution_WAGE_Jan2024
#| fig-cap: "Distribution of Wages (January 1–15, 2024)."
#| echo: FALSE
knitr::include_graphics(here("results","figures","distribution_WAGE_Jan2024.png"))
```

```{r}
#| label: tbl-combined_accuracy1
#| tbl-cap: "Model Fit for Wage Based on Job Title."
#| echo: FALSE
table1 = readRDS(here("results","tables","combined_accuracy1.rds"))
knitr::kable(table1)
```

```{r}
#| label: tbl-combined_accuracy2
#| tbl-cap: "Model Fit for Wage Based on Employer Locations."
#| echo: FALSE
table2 = readRDS(here("results","tables","combined_accuracy2.rds"))
knitr::kable(table2)
```

```{r}
#| label: tbl-combined_accuracy3
#| tbl-cap: "Model Fit for Wage Based on Worksite Locations."
#| echo: FALSE
table3 = readRDS(here("results","tables","combined_accuracy3.rds"))
knitr::kable(table3)
```

```{r}
#| label: tbl-combined_accuracy4
#| tbl-cap: "Model Fit for Wage Based on Employer."
#| echo: FALSE
table4 = readRDS(here("results","tables","combined_accuracy4.rds"))
knitr::kable(table4)
```

```{r}
#| label: tbl-combined_accuracy_all
#| tbl-cap: "Model Fit for Wage Based on Job Title, Employer Location, Worksite Location, and Employer."
#| echo: FALSE
table5 = readRDS(here("results","tables","combined_accuracy_all.rds"))
knitr::kable(table5)
```

## Full analysis

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

## Strengths and Limitations

## Conclusions

{{< pagebreak >}}

# References