### 2025-04-02 products/manuscript/supplement/Supplementary-Material.qmd

---
title: "Supplement to Yufei Wu's MADA Project"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Overview

You can find all the supplemetary materials (figures and tables) in the supplement.

# Additional results
## Exploratory/Descriptive analysis

@tbl-sum_prev_wage shows a summary table for prevailing wage.

```{r}
#| label: tbl-sum_prev_wage
#| tbl-cap: "Summary Table for Prevailing Wage."
#| echo: FALSE
table1 = readRDS(here("results","tables","sum_prev_wage.rds"))
knitr::kable(table1)
```

@fig-bottom_wage_by_soc_title shows a bar plot of bottom 10 job titles by lowest median wage. @fig-bottom_wage_by_employer shows a bar plot of bottom 10 employers by lowest median wage. @fig-bottom_wage_by_state shows a bar plot of bottom 10 states by lowest median wage.

```{r}
#| label: fig-bottom_wage_by_soc_title
#| fig-cap: "Bottom 10 Job Titles by Lowest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","bottom_wage_by_soc_title.png"))
```

```{r}
#| label: fig-bottom_wage_by_employer
#| fig-cap: "Bottom 10 Employers by Lowest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","bottom_wage_by_employer.png"))
```

```{r}
#| label: fig-bottom_wage_by_state
#| fig-cap: "Bottom 10 States by Lowest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","bottom_wage_by_state.png"))
```

## Model Fitting

@tbl-rf_combined_accuracy_cv_default to @tbl-combined_accuracy_cv_svm show the model accuracy in both training and test data of the three models: Random Forest, XGBoost, and SVM.

```{r}
#| label: tbl-rf_combined_accuracy_cv_default
#| tbl-cap: "Accuracy of Random Forest."
#| echo: FALSE
table2 = readRDS(here("results","tables","rf_combined_accuracy_cv_default.rds"))
knitr::kable(table2)
```

```{r}
#| label: tbl-combined_accuracy_cv_xgb_reg
#| tbl-cap: "Accuracy of XGBoost."
#| echo: FALSE
table3 = readRDS(here("results","tables","combined_accuracy_cv_xgb_reg.rds"))
knitr::kable(table3)
```

```{r}
#| label: tbl-combined_accuracy_cv_svm
#| tbl-cap: "Accuracy of SVM."
#| echo: FALSE
table4 = readRDS(here("results","tables","combined_accuracy_cv_svm.rds"))
knitr::kable(table4)
```

{{< pagebreak >}}

# Discussion

{{< pagebreak >}}